Handling Recursion in Generic Programming Using Closed Type Families

### 2018-05-10 lncs-paper.tex
Typos and some wibble

The following exploits the idea of \emph{pattern matching}. As before, we consider two cases. In the first case, when the position we are matching on is not recursive, we only require it to be an  instance of \texttt{Show}, and invoke its \texttt{show} function. Whereas in the case of the recursive position, we surround it with parentheses and apply our generic function \texttt{gshow}. Thus, by means of the type family for equality, we model a form of pattern matching on the types again:
\texttt{
\vs
\indent\textbf{class} CaseShow (eq \ann Bool) (a \ann *) (b \ann *) \textbf{where}\\
\indent\s\s caseShow' \ann b \ra String
\vs
\indent\textbf{instance} Show\s\s b \Ra CaseShow 'False a b \textbf{where}\\
\indent\s\s caseShow'\s\s\s = show\\
\indent\textbf{instance} GShow a \Ra CaseShow 'True\s\s a a \textbf{where}\\
\indent\s\s caseShow' t = "(" ++ gshow t ++ ")"
\vs
}
We provide a synonym for the \texttt{CaseShow (Equal a b) a b} instance, which we call \texttt{CaseRecShow}, as before with \texttt{ProofEq}; likewise the synonym for the matching function:
\texttt{
\vs
\indent caseShow \ann \fa a b. CaseRecShow a b \Ra b \ra String\\
\indent caseShow t = caseShow' @(Equal a b) @a t
\vspace{0.2cm}
}

The resulting function \texttt{gshow} is a subject of a number of constraints abbreviated by \texttt{GShow} synonym:
\texttt{
\vs
\indent\textbf{type} GShow a = (Generic a, HasDatatypeInfo a,\\
\indent\ind\ind\ind\ind All2 (CaseRecShow a) (Code a))
\vs
\indent gshow \ann \fa a. GShow a \Ra a \ra String
\vs
}
The function \texttt{gshow} employs metadata provided by \textsf{generics-sop}'s class \texttt{Has\-Datatype\-Info} to show the names of a datatype constructor and its record fields. The \textsf{generics-sop} library is able to derive this metadata automatically.
The function is also constrained by \texttt{CaseRecShow} with the \texttt{All2} combinator that is an analogue of \texttt{All} for a list of lists of types.

@@

The zipper is a data structure that enables efficient navigation and editing within the tree-like structure of a datatype. It represents a current location in that structure, storing a tree node, a \emph{focus}, along with its context. Having a zipper focused on a recursive knot in a structure, we may produce a new location by moving the focus up, down, left, or right. On the way, we can update the nodes. Entering and laving the navigation usually need a special care.

@@

The \texttt{update} function modifies the focal subtree with a given constrained function. The type in focus is existentially quantified inside \texttt{Loc} and should satisfy constraint \texttt{c}. The structure of \texttt{Loc} (shown in Section~\ref{ss:locs}) guarantees that the constraint holds for all types in the family \texttt{fam} and, therefore, for all recursive nodes that can be in focus, hence \texttt{update} can always be applied.

@@

We implement the \texttt{In} constraint by means of a type family \texttt{InFam} exactly along the lines of \texttt{Equal} family defined in the beginning of Section~\ref{sec:handling-recursion}. 
%~ The following checks a type for membership of a type-level list that we use to express a family:
%~ \texttt{
%~ \vs
%~ \indent type family InFam (a \ann k) (fam \ann [k]) \ann Bool where\\
%~ \indent\s\s InFam a (a ': fam) = 'True\\
%~ \indent\s\s InFam a (x ': fam) = InFam a fam\\
%~ \indent\s\s InFam a '[]\ind\ind = 'False
%~ \vs
%~ }
%~ The \texttt{In} constraint is defined as
\texttt{
\vs
\indent type In a fam = InFam a fam \tteq 'True\\
\vs
}
The definition of \texttt{InFam} is omitted as a boring one.

One last missing piece for managing focuses is the class \texttt{ProofIn}. it provides a proof of membership of a focus type to a family. Again, this generalizes the proof of type equality from Section~\ref{sec:handling-recursion}. The definitions of \texttt{ProofIn} and an auxiliary class \texttt{ProofFam} are given in Figure~\ref{fig:proof-in}. 
%The constraints on the second \texttt{ProofFam} instance, which repeat the same ones from the \texttt{Family} constructor, constitute the proof.
% this is not clear and, I think, can be omitted

@@

The idea similar to SOP has been proposed by Kiselyov~et~al.~\cite{Kiselyov2004} in their \textsf{HList} library for strongly typed heterogeneous collections. In the paper, the authors also discuss problems connected with overlap, which they use for access operations. Another Haskell extension, functional dependencies, is applied to restrict overlap in type equality there, which resembles our solution.

>>>>

The following exploits the idea of \emph{pattern matching}. As before, we consider two cases. In the first case, when the position we are matching on is not recursive, we only require it to be an  instance of \texttt{Show}, and invoke its \texttt{show} function. Whereas in the case of the recursive position, we surround it with parentheses and apply our generic function \texttt{gshow}. Thus, by means of the type family for equality, we model a form of pattern matching on the types again:
\texttt{
\vs
\indent\textbf{class} CaseShow (eq \ann Bool) (a \ann *) (b \ann *) \textbf{where}\\
\indent\s\s caseShow' \ann b \ra String
\vs
\indent\textbf{instance} Show\s\s b \Ra CaseShow 'False a b \textbf{where}\\
\indent\s\s caseShow'\s\s\s = show\\
\indent\textbf{instance} GShow a \Ra CaseShow 'True\s\s a a \textbf{where}\\
\indent\s\s caseShow' t = "(" ++ gshow t ++ ")"
\vs
}
We provide a synonym for the \texttt{CaseShow (Equal a b) a b} instance, which we call \texttt{CaseRecShow}, as before with \texttt{ProofEq}; likewise a synonym for the matching function:
\texttt{
\vs
\indent caseShow \ann \fa a b. CaseRecShow a b \Ra b \ra String\\
\indent caseShow t = caseShow' @(Equal a b) @a t
\vspace{0.2cm}
}

The resulting function \texttt{gshow} is a subject of a number of constraints abbreviated by a \texttt{GShow} synonym:
\texttt{
\vs
\indent\textbf{type} GShow a = (Generic a, HasDatatypeInfo a,\\
\indent\ind\ind\ind\ind All2 (CaseRecShow a) (Code a))
\vs
\indent gshow \ann \fa a. GShow a \Ra a \ra String
\vs
}
The function \texttt{gshow} employs metadata provided by \textsf{generics-sop}'s class \texttt{Has\-Datatype\-Info} to show the names of a datatype constructor and its record fields. The \textsf{generics-sop} library is able to derive this metadata automatically.
The function is also constrained by \texttt{CaseRecShow} with the \texttt{All2} combinator that is an analogue of \texttt{All} for a list of lists of types.

@@

The zipper is a data structure that enables efficient navigation and editing within the tree-like structure of a datatype. It represents a current location in that structure, storing a tree node, a \emph{focus}, along with its context. Having a zipper focused on a recursive knot in a structure, we may produce a new location by moving the focus up, down, left, or right. On the way, we can update the nodes. Entering and leaving the navigation usually need a special care.

@@

The \texttt{update} function modifies the focal subtree with a given constrained function. The type in focus is existentially quantified inside \texttt{Loc} and should satisfy the constraint \texttt{c}. The structure of \texttt{Loc} (shown in Section~\ref{ss:locs}) guarantees that the constraint holds for all types in the family \texttt{fam} and, therefore, for all recursive nodes that can be in focus, hence \texttt{update} can always be applied.

@@

We implement the \texttt{In} constraint by means of a type family \texttt{InFam} exactly along the lines of the \texttt{Equal} type family defined in the beginning of Section~\ref{sec:handling-recursion}. 
%~ The following checks a type for membership of a type-level list that we use to express a family:
%~ \texttt{
%~ \vs
%~ \indent type family InFam (a \ann k) (fam \ann [k]) \ann Bool where\\
%~ \indent\s\s InFam a (a ': fam) = 'True\\
%~ \indent\s\s InFam a (x ': fam) = InFam a fam\\
%~ \indent\s\s InFam a '[]\ind\ind = 'False
%~ \vs
%~ }
%~ The \texttt{In} constraint is defined as
\texttt{
\vs
\indent type In a fam = InFam a fam \tteq 'True\\
\vs
}
The definition of \texttt{InFam} is omitted as a boring one.

One last missing piece for managing focuses is the class \texttt{ProofIn}. It provides a proof of membership of a focus type to a family. Again, this generalizes the proof of type equality from Section~\ref{sec:handling-recursion}. The definitions of \texttt{ProofIn} and an auxiliary class \texttt{ProofFam} are given in Figure~\ref{fig:proof-in}. 
%The constraints on the second \texttt{ProofFam} instance, which repeat the same ones from the \texttt{Family} constructor, constitute the proof.
% this is not clear and, I think, can be omitted

@@

The idea similar to SOP has been proposed by Kiselyov~et~al.~\cite{Kiselyov2004} in their \textsf{HList} library for strongly typed heterogeneous collections. In the paper, the authors also discuss problems connected with overlap, which they use for access operations. Another Haskell extension, functional dependencies, is applied to restrict overlap by introducing a class for type equality there, which resembles our solution.
